{\fontfamily{ptm}\selectfont Nesterov Type Momentum Methods}

### 2024-07-07 Notebook LaTEX/Momentum Accelerations/main.tex
>w< more small updates to the notes

            \begin{definition}[Ahn Sra 6.24]\label{def:agg_ppm}
                With $f$ $\mu \ge 0$ strongly convex and $L$-Lipschitz smooth, the generic PPM form is formulated for strictly positive stepsizes $\tilde \eta_i,\eta_i$: 
                \begin{align*}
                    x_{t + 1} &= \argmin_{x} \left\lbrace
                        l_f(x; y_t) 
                        + 
                        \frac{\mu}{2}\Vert x - y_t\Vert^2
                        + 
                        \frac{1}{2\tilde \eta_{t + 1}} \Vert x - x_t\Vert^2
                    \right\rbrace, 
                    \\
                    y_{t + 1} &= \argmin_{x} 
                    \left\lbrace
                        l_f(x; y_t) + \frac{L}{2}\Vert x - y_t\Vert^2 + 
                        \frac{1}{2\eta_{t + 1}}\Vert x - x_{t + 1}\Vert^2
                    \right\rbrace. 
                \end{align*}
            \end{definition}
            \begin{remark}
                This algorithm is the same as algorithm (6.24) described by Ahn and Sra \cite{ahn_understanding_2022}. 
                Observe that by setting $\mu = 0, \tilde \eta_{t + 1} = \eta_{t + 1}$ this recovers algorithm (4.8) described in Ahn and Sra \cite{ahn_understanding_2022}. 
            \end{remark}

            \begin{definition}[AG Proximal Gradient PPM Generic Form]
            \label{def:ag_prox_grad_ppm}
                Let $h=f + g$ be the sum of convex function $g$ and convex differentiable $f$ with $L$-Lipschitz gradient. 
                Define the proximal gradient and gradient mapping operator operator: 
                $$
                \begin{aligned}
                    \mathcal T_L^{g, f}(x) &:=  
                    \hprox_{L^{-1}g}\left(x - L^{-1}\nabla f(x)\right)
                    \\
                    &= 
                    \argmin_{u}
                    \left\lbrace
                        g(u) + f(x) + \langle \nabla f(x), u - x\rangle
                        + 
                        \frac{L}{2}\Vert u - x\Vert^2
                    \right\rbrace,
                    \\
                    \mathcal G_L^{g, f}(x) &= 
                    L\left(x - \mathcal T_L^{g, f}(x)\right). 
                \end{aligned}
                $$
                Omitting the superscript $f, g$ on $\mathcal P, \mathcal G$ for simplicity since it's clear in the context. 
                Define the linear lower bounding function for $f$ at $y$, for all $x$: 
                $$
                \begin{aligned}
                    l_h(x; y) &= h(\mathcal T_L y) + \langle \mathcal G_L(y), x - y \rangle \le f(x), 
                \end{aligned}
                $$
                With that we define the algorithm:
                $$
                \begin{aligned}
                    x_{t + 1} &= \argmin_{x} \left\lbrace
                        l_h(x; y_t) + \frac{1}{2\tilde \eta_{t + 1}} 
                        \Vert x - x_t\Vert^2
                    \right\rbrace,
                    \\
                    y_{t + 1}&= 
                    \argmin_{x}
                    \left\lbrace
                        g(x) + l_f(x; y_t) + \frac{L}{2}\Vert x - y_t\Vert^2 + 
                        \frac{1}{2\eta_{t + 1}} \Vert x - x_{t + 1}\Vert^2
                    \right\rbrace.
                \end{aligned}
                $$
            \end{definition}

@@

            \begin{definition}[Tri-points Generic Form]\label{def:agg_tri}
                With $f$ be $L$-Lipschitz \\ 
                smooth and convex, choose any $y_0 = x_0=z_0$ and a non-negative sequence $\eta_t, \tilde\eta_t$, the algorithm admits form: 
                \begin{align*}
                    x_{t + 1} &= x_t - \tilde \eta_{t + 1} \nabla f(y_t) 
                    \\
                    z_{t + 1} &= y_t - L^{-1} \nabla f(y_t) 
                    \\
                    y_{t + 1} &= 
                    (1 + L\eta_{t + 1})^{-1}
                    (
                    x_{t + 1} + L\eta_{t + 1}z_{t + 1}
                    ). 
                \end{align*}
            \end{definition}
            \begin{remark}
                The parameter $\eta_t, \tilde\eta_{t + 1}$ are exactly the same as Ahn, Sra \cite[(6.24)]{ahn_understanding_2022}, but with the choice of $\mu = 0$. 
                This form is not explictly stated by Ahn and Sra but it's trivial to realize its presence regardless. 
                They are the same parameters for both algorithms and they are equivalent. 
                This will be proved in the coming parts. 
            \end{remark}
            \begin{definition}[Strongly Convex Tri-points Generic Form]
                \quad 
                With $f$ being $L$-Lipschitz smooth and $\mu > 0$ strongly convex then the Strongly convex Tri-Points generic form has updates of for iterates
                \begin{align*}
                    x_{t + 1} &= (1 + \mu\tilde \eta_{t + 1})^{-1}(x_t + \mu\tilde\eta_{t + 1}y_t^+), 
                    \\
                    z_{t + 1} &= y_t - L^{-1}\nabla f(y_t), 
                    \\
                    y_{t + 1} &= (1 + L\eta_{t + 1})^{-1}(x_{t + 1} + L\eta_{t + 1}z_{t +1}). 
                \end{align*}
            \end{definition}

@@

            \begin{definition}[AG Ryu 12.1]
            \label{def:ag_ryu_12.1}
                State in chapter 12 in Ryu's writing \cite{ryu_large-scale_2022} is the following variant of accelerated gradient algorithm: 
                \begin{align*}
                    z_{t + 1} &= y_k + L^{-1}\nabla f(y_t), 
                    \\
                    y_{t + 1} &= z_{t + 1} + \frac{t - 1}{t + 2}\left(
                        z_{k + 1} - z_k
                    \right). 
                \end{align*}
                An equivalent form of the above is also provided by Ryu: 
                \begin{align*}
                    z_{t + 1} &= y_t + L^{-1}\nabla f(y_t), 
                    \\
                    x_{t + 1} &= x_t + (t + 1)(2L)^{-1}\nabla f(y_t), 
                    \\
                    y_{t + 1} &= \left(
                        1 - \frac{2}{t + 2} 
                    \right)z_{t + 1} + 
                    \left(
                        \frac{1}{t + 2}
                    \right)x_{t + 1}. 
                \end{align*}
            \end{definition}
            \begin{observation}
                The base case requires cares since it's not unique, there are more representations  depending on how the base case is define, 
                Observe that the second equivalent form is an example of \hyperref[def:agg_tri]{definition \ref*{def:agg_tri}}. 
                \par\noindent
                It is stated in Ryu's writing that the base case is $x_0 = y_0 = z_0$, 
                \textcolor{purple}{we think this is not the only options.} 
                We remind the reader that there is not an obvious choice of $t\in \mathbb Z$ such that $x_t = y_t = z_t$ where the above algorithm remains consistent for all $t \ge -1$. 
                The choice of base case presented in Ryu's writing is a choice and not a necessity. 
                We state that one of the base case scenario that only requires knowing $y_1$.

                \par\noindent 
                We need to observe the case where $t = 0$ to understand the base case of the algorithm. 
                At each iteration $t$, Given $(x_t, y_t)$, the above formula maps to $(x_{t + 1}, y_{t + 1}, z_{t + 1})$. 
                When $t= 0$, it has $y_1 = x_1$. 
                Therefore, knowing only either $y_1$, or $x_1$ can initiate the algorithm. 
                This is an alternative scenario of the base case. 
                If, we wish to write $y_0 = x_0$ at $t = 0$ for the base case instead, then it creates the following algorithm 
                \begin{align*}
                    y_{t} &= \left(
                    1 - \frac{2}{t + 2} 
                    \right)z_{t} + 
                    \left(
                        \frac{1}{t + 2}
                    \right)x_{t}, 
                    \\
                    z_{t + 1} &= y_t + L^{-1}\nabla f(y_t), 
                    \\
                    x_{t + 1} &= x_t + (t + 1)(2L)^{-1}\nabla f(y_t). 
                \end{align*}
                It has a different representations due to a specific choice of the base case. 
                \par\noindent
                Finally, by the observation that given $(z_t, y_t)$ for any $t\ge -1$, we can solve for $x_t$ \newline via $y_t = (1 - 2/(1 + t))z_t + 1/(t + 1)x_t$, obtaining $(x_t, y_t, z_t)$, hence the algorithm can be reduced to a representations with only $(z_t, y_t)$ and an updates using only $(z_t, y_t)$ to $(z_{t + 1}, y_{t + 1})$. 
                That is the momentum form presented above. 
            \end{observation}
            \begin{remark}
                We advise the reader to read the Biliography notes by Ryu \cite[chapter 12]{ryu_large-scale_2022} of his book. 
                It goes over the full context and history for this particular variant of gradient acceleration method. 
                As the time of composing this notes, the writer is still reading on the topic of accelerated gradient.     
            \end{remark}

            \begin{definition}[Form Zero]
            \label{def:ag_form_zero}

>>>>

            \begin{definition}[Ahn Sra 6.24]\label{def:agg_ppm}
                With $f$ $\mu \ge 0$ strongly convex and $L$-Lipschitz smooth, the generic PPM form is formulated for strictly positive stepsizes $\tilde \eta_i,\eta_i$: 
                \begin{align*}
                    x_{t + 1} &= \argmin_{x} \left\lbrace
                        l_f(x; y_t) 
                        + 
                        \frac{\mu}{2}\Vert x - y_t\Vert^2
                        + 
                        \frac{1}{2\tilde \eta_{t + 1}} \Vert x - x_t\Vert^2
                    \right\rbrace, 
                    \\
                    y_{t + 1} &= \argmin_{x} 
                    \left\lbrace
                        l_f(x; y_t) + \frac{L}{2}\Vert x - y_t\Vert^2 + 
                        \frac{1}{2\eta_{t + 1}}\Vert x - x_{t + 1}\Vert^2
                    \right\rbrace. 
                \end{align*}
            \end{definition}
            \begin{remark}
                This algorithm is the same as algorithm (6.24) described by Ahn and Sra \cite{ahn_understanding_2022}. 
                Observe that by setting $\mu = 0, \tilde \eta_{t + 1} = \eta_{t + 1}$ this recovers algorithm (4.8) described in Ahn and Sra \cite{ahn_understanding_2022}. 
                Finally, Ahn, Sra 6.24 has the same form as Nesterov 2.2.7. 
                See 
                \hyperref[prop:Nes2.2.7_via_ahn_sra_6.24]{proposition \ref*{prop:Nes2.2.7_via_ahn_sra_6.24}} for a demonstration. 
            \end{remark}

            \begin{definition}[AG Proximal Gradient PPM Generic Form]
            \label{def:ag_prox_grad_ppm}
                Let $h=f + g$ be the sum of convex function $g$ and convex differentiable $f$ with $L$-Lipschitz gradient. 
                Define the proximal gradient and gradient mapping operator operator: 
                $$
                \begin{aligned}
                    \mathcal T_L^{g, f}(x) &:=  
                    \hprox_{L^{-1}g}\left(x - L^{-1}\nabla f(x)\right)
                    \\
                    &= 
                    \argmin_{u}
                    \left\lbrace
                        g(u) + f(x) + \langle \nabla f(x), u - x\rangle
                        + 
                        \frac{L}{2}\Vert u - x\Vert^2
                    \right\rbrace,
                    \\
                    \mathcal G_L^{g, f}(x) &= 
                    L\left(x - \mathcal T_L^{g, f}(x)\right). 
                \end{aligned}
                $$
                Omitting the superscript $f, g$ on $\mathcal P, \mathcal G$ for simplicity since it's clear in the context. 
                Define the linear lower bounding function for $f$ at $y$, for all $x$: 
                $$
                \begin{aligned}
                    l_h(x; y) &= h(\mathcal T_L y) + \langle \mathcal G_L(y), x - y \rangle \le f(x), 
                \end{aligned}
                $$
                With that we define the algorithm:
                $$
                \begin{aligned}
                    x_{t + 1} &= \argmin_{x} \left\lbrace
                        l_h(x; y_t) + \frac{1}{2\tilde \eta_{t + 1}} 
                        \Vert x - x_t\Vert^2
                    \right\rbrace,
                    \\
                    y_{t + 1}&= 
                    \argmin_{x}
                    \left\lbrace
                        g(x) + l_f(x; y_t) + \frac{L}{2}\Vert x - y_t\Vert^2 + 
                        \frac{1}{2\eta_{t + 1}} \Vert x - x_{t + 1}\Vert^2
                    \right\rbrace.
                \end{aligned}
                $$
            \end{definition}
            \begin{observation}
                With $g \equiv 0$, the above definition reduces to 
                \hyperref[def:agg_ppm]{definition \ref*{def:agg_ppm}}
                where $\mu = 0$. 
            \end{observation}

@@

            \begin{definition}[Tri-points Generic Form]\label{def:agg_tri}
                With $f$ be $L$-Lipschitz \\ 
                smooth and convex, choose any $y_0 = x_0=z_0$ and a non-negative sequence $\eta_t, \tilde\eta_t$, the algorithm admits form: 
                \begin{align*}
                    x_{t + 1} &= x_t - \tilde \eta_{t + 1} \nabla f(y_t) 
                    \\
                    z_{t + 1} &= y_t - L^{-1} \nabla f(y_t) 
                    \\
                    y_{t + 1} &= 
                    (1 + L\eta_{t + 1})^{-1}
                    (
                    x_{t + 1} + L\eta_{t + 1}z_{t + 1}
                    ). 
                \end{align*}
            \end{definition}
            \begin{remark}
                The parameter $\eta_t, \tilde\eta_{t + 1}$ are exactly the same as Ahn, Sra \cite[(6.24)]{ahn_understanding_2022}, but with the choice of $\mu = 0$. 
                This form is not explictly stated by Ahn and Sra but it's trivial to realize its presence regardless. 
                They are the same parameters for both algorithms and they are equivalent. 
                For a demonstration, see 
                \hyperref[prop:tri_form_via_ppm]{proposition \ref*{prop:tri_form_via_ppm}}. 
            \end{remark}

            \begin{definition}[Strongly Convex Tri-points Generic Form]
                \quad \\
                With $f$ being $L$-Lipschitz smooth and $\mu > 0$ strongly convex then the Strongly convex Tri-Points generic form has updates of for iterates
                \begin{align*}
                    x_{t + 1} &= (1 + \mu\tilde \eta_{t + 1})^{-1}(x_t + \mu\tilde\eta_{t + 1}y_t^+), 
                    \\
                    z_{t + 1} &= y_t - L^{-1}\nabla f(y_t), 
                    \\
                    y_{t + 1} &= (1 + L\eta_{t + 1})^{-1}(x_{t + 1} + L\eta_{t + 1}z_{t +1}). 
                \end{align*}
            \end{definition}
            \begin{remark}
                This algorithm is equivalent to 
                \hyperref[def:agg_ppm]{definition \ref*{def:agg_ppm}}. 
                See \ref*{prop:tri_scvx_from_ahn_sra_6.24} for a demonstration. 
            \end{remark}

@@

            \begin{definition}[AG Ryu 12.1]
            \label{def:ag_ryu_12.1}
                State in chapter 12 in Ryu's writing \cite{ryu_large-scale_2022} is the following variant of accelerated gradient algorithm: 
                \begin{align*}
                    z_{t + 1} &= y_k + L^{-1}\nabla f(y_t), 
                    \\
                    y_{t + 1} &= z_{t + 1} + \frac{t - 1}{t + 2}\left(
                        z_{k + 1} - z_k
                    \right). 
                \end{align*}
                An equivalent form of the above is also provided by Ryu: 
                \begin{align*}
                    z_{t + 1} &= y_t + L^{-1}\nabla f(y_t), 
                    \\
                    x_{t + 1} &= x_t + (t + 1)(2L)^{-1}\nabla f(y_t), 
                    \\
                    y_{t + 1} &= \left(
                        1 - \frac{2}{t + 2} 
                    \right)z_{t + 1} + 
                    \left(
                        \frac{2}{t + 2}
                    \right)x_{t + 1}. 
                \end{align*}
            \end{definition}
            \begin{observation}
                The base case requires cares since it's not unique, there are more representations  depending on how the base case is define, 
                Observe that the second equivalent form is an example of \hyperref[def:agg_tri]{definition \ref*{def:agg_tri}}. 
                \par\noindent
                It is stated in Ryu's writing that the base case is $x_0 = y_0 = z_0$, 
                we think this is not the only options. 
                We remind the reader that there is not an obvious choice of $t\in \mathbb Z$ such that $x_t = y_t = z_t$ where the above algorithm remains consistent for all $t \ge -1$. 
                The choice of base case presented in Ryu's writing is a choice and not a necessity. 
                We state that one of the base case scenario that only requires knowing $y_1$.

                \par\noindent 
                We need to observe the case where $t = 0$ to understand the base case of the algorithm. 
                At each iteration $t$, Given $(x_t, y_t)$, the above formula maps to $(x_{t + 1}, y_{t + 1}, z_{t + 1})$. 
                When $t= 0$, it has $y_1 = x_1$. 
                Therefore, knowing only either $y_1$, or $x_1$ can initiate the algorithm. 
                This is an alternative scenario of the base case. 
                If, we wish to write $y_0 = x_0$ at $t = 0$ for the base case instead, then it creates the following algorithm 
                \begin{align*}
                    y_{t} &= \left(
                    1 - \frac{2}{t + 2} 
                    \right)z_{t} + 
                    \left(
                        \frac{2}{t + 2}
                    \right)x_{t}, 
                    \\
                    z_{t + 1} &= y_t + L^{-1}\nabla f(y_t), 
                    \\
                    x_{t + 1} &= x_t + (t + 1)(2L)^{-1}\nabla f(y_t). 
                \end{align*}
                It has a different representations due to a specific choice of the base case. 
                \par\noindent
                Finally, by the observation that given $(z_t, y_t)$ for any $t\ge -1$, we can solve for $x_t$ \newline via $y_t = (1 - 2/(1 + t))z_t + 1/(t + 1)x_t$, obtaining $(x_t, y_t, z_t)$, hence the algorithm can be reduced to a representations with only $(z_t, y_t)$ and an updates using only $(z_t, y_t)$ to $(z_{t + 1}, y_{t + 1})$. 
                That is the momentum form presented above. 
            \end{observation}
            \begin{remark}
                We advise the reader to read the Biliography notes by Ryu \cite[chapter 12]{ryu_large-scale_2022} of his book. 
                It goes over the full context and history for this particular variant of gradient acceleration method. 
                As the time of composing this notes, the writer is still reading on the topic of accelerated gradient.     
            \end{remark}

            \begin{definition}[Proximal Gradient Tri-Points Form E]
            \label{def:ag_tri_pt_form_E}
                Form E is a special case of 
                \hyperref[def:ag_prox_grad_tri_pt]{definition \ref*{def:ag_prox_grad_tri_pt}}
                with the condition that $\tilde \eta_t = \eta_t$ for all $t \in \mathbb N$. 
                As a consequence, the algorithm has the following updates for iterates $(y_t, x_{t + 1}, z_{t + 1})$. 
                Giving us: 
                \begin{align*}
                    y_t &= (1 + L\eta_t)^{-1}(x_t + L\eta_t z_t)
                    \\
                    x_{t + 1} &= x_t - \eta_t \mathcal G_L(y_t)
                    \\
                    z_{t + 1} &= y_t - L^{-1}\mathcal G_L(y_t).
                \end{align*}
            \end{definition}
            \begin{remark}
